Decoding the Impact: Statistical Analysis of Factors Contributing to NFL Injuries

### 2024-04-23 paper.qmd
more changes

---
title: "Decoding the Impact: Statistical Analysis of Factors Contributing to NFL Injuries"
author: "Vincent Xie"
toc: false
number-sections: false
execute:
    echo: false
highlight-style: pygments
format: 
  html: 
    code-fold: true
    html-math-method: katex
  pdf: 
    fontsize: 12pt
    geometry: 
      - top=30mm
      - left=20mm
header-includes:
  - \usepackage{setspace}
  - \doublespacing
##  docx: default
---
```{python}
# Packages
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
import statsmodels.api as sm
from scipy.stats import chi2_contingency
```

@@

```{python}
# Load the datasets
injury = pd.read_csv('data/InjuryRecord.csv')
plays = pd.read_csv('data/PlayList.csv')
```

@@

```{python}
# Bar chart for frequency of stadium types in PlayList
stadium_freq = plays['StadiumType'].value_counts()
plt.figure(figsize=(8, 4))
stadium_freq.plot(kind='bar', color='lightgreen')
plt.title('Frequency of Stadium Types')
plt.xlabel('Stadium Type')
plt.ylabel('Frequency')
plt.show()
```

@@

@@

# Bar chart for frequency of stadium types in PlayList
stadium_freq = plays['StadiumType'].value_counts()
plt.figure(figsize=(8, 4))
stadium_freq.plot(kind='bar', color='lightgreen')
plt.title('Frequency of Stadium Types')
plt.xlabel('Stadium Type')
plt.ylabel('Frequency')
plt.show()
```

@@

```{python}
# Bar chart for frequency of each injury type in InjuryRecord
injury_freq = injury['BodyPart'].value_counts()
plt.figure(figsize=(5, 2))
injury_freq.plot(kind='bar', color='skyblue')
plt.title('Frequency of Each Injury')
plt.xlabel('Body Part')
plt.ylabel('Frequency')
plt.show()

# Merge injury_record and playlist on PlayerKey
merged_data = pd.merge(injury, plays, on='PlayKey', how='inner')

@@

```{python}
# Merge injury_record and playlist on PlayerKey
merged_data = pd.merge(injury, plays, on='PlayKey', how='inner')

@@

```{python}
# Create a set of unique keys from InjuryRecord which indicates injuries
injury_keys = set(injury['PlayKey'])

@@

>>>>

---
title: "Decoding the Impact: Statistical Analysis of Factors Contributing to NFL Injuries"
author: "Vincent Xie"
toc: false
number-sections: false
execute:
    echo: false
highlight-style: pygments
format: 
  html: 
    code-fold: true
    html-math-method: katex
  pdf: 
    fontsize: 12pt
    geometry: 
      - top=30mm
      - left=20mm
header-includes:
  - \usepackage{setspace}
  - \doublespacing
##  docx: default
---
```{python}
# Packages
import pandas as pd
import matplotlib.pyplot as plt
import seaborn as sns
import statsmodels.api as sm
from scipy.stats import chi2_contingency
import numpy as np
```

@@

```{python}
# Load the datasets
injury = pd.read_csv('data/InjuryRecord.csv')
plays = pd.read_csv('data/PlayList.csv')
new_injury = pd.read_csv('data/mod_InjuryRecord.csv')
```

@@

```{python}
# Bar chart for frequency of stadium types in PlayList
stadium_freq = plays['StadiumType'].value_counts()
plt.figure(figsize=(5, 2))
stadium_freq.plot(kind='bar', color='lightgreen')
plt.title('Frequency of Stadium Types')
plt.xlabel('Stadium Type')
plt.ylabel('Frequency')
plt.show()
```

@@

# Save the modified dataframe if needed
plays.to_csv('data/mod_PlayList.csv', index=False)

@@

# Bar chart for frequency of stadium types in PlayList
stadium_freq = plays['StadiumType'].value_counts()
plt.figure(figsize=(6, 2))
stadium_freq.plot(kind='bar', color='lightgreen')
plt.title('Frequency of Stadium Types')
plt.xlabel('Stadium Type')
plt.ylabel('Frequency')
plt.show()
```

We can also
```{python}
# Define the temperature bins and labels
bins = [-float('inf'), 50, 60, 70, 80, float('inf')]
labels = ['Cold', 'Cool', 'Mild', 'Warm', 'Hot']

# Replace the temperature values with categorical labels directly
plays['Temperature'] = pd.cut(plays['Temperature'], bins=bins, labels=labels)

# Save the modified dataframe if needed
plays.to_csv('data/mod_PlayList.csv', index=False)
```

```{python}
# Mapping function for weather descriptions
def map_weather(weather):
    if pd.isna(weather):
        return 'Unknown'
    weather = weather.lower()
    if any(word in weather for word in ['rain', 'showers', 'rainy']):
        return 'Rain'
    elif any(word in weather for word in ['overcast', 'cloudy']):
        return 'Overcast'
    elif any(word in weather for word in ['clear', 'sunny']):
        return 'Clear'
    elif 'snow' in weather:
        return 'Snow'
    elif any(word in weather for word in ['indoor', 'indoors', 'controlled climate', 'n/a']):
        return 'None'
    return 'Unknown'

# Apply the mapping function to the Weather column
plays['WeatherCategory'] = plays['Weather'].apply(map_weather)

# Save the modified dataframe if needed
plays.to_csv('data/mod_PlayList.csv', index=False)

# Verify the changes by displaying the value counts
plays['WeatherCategory'].value_counts()
```

@@

```{python}
# Bar chart for frequency of each injury type in InjuryRecord
injury_freq = new_injury['BodyPart'].value_counts()
plt.figure(figsize=(5, 2))
injury_freq.plot(kind='bar', color='skyblue')
plt.title('Frequency of Each Injury')
plt.xlabel('Body Part')
plt.ylabel('Frequency')
plt.show()

# Merge injury_record and playlist on PlayerKey
merged_data = pd.merge(new_injury, plays, on='PlayKey', how='inner')

@@

```{python}
# Merge injury_record and playlist on PlayerKey
merged_data = pd.merge(new_injury, plays, on='PlayKey', how='inner')

@@

```{python}
# Create a set of unique keys from InjuryRecord which indicates injuries
injury_keys = set(new_injury['PlayKey'])

@@

```{}
# If 'FieldType' is not already coded as 0/1, convert it to numeric codes
plays['FieldType'] = plays['FieldType'].astype('category').cat.codes

# Filter the DataFrame to include only cases where an injury occurred
injury_cases = plays[plays['InjuryOccurred'] == 1]

# The independent variable should be 'FieldType', and we add an intercept to it
X = injury_cases[['FieldType']]  # Assuming 'FieldType' is already coded as 0/1
X = sm.add_constant(X)  # This adds the intercept

# The dependent variable is 'InjuryOccurred' (though in this subset, it will always be 1)
y = injury_cases['InjuryOccurred']

# Perform logistic regression
logit_model = sm.Logit(y, X)
result = logit_model.fit(maxiter=100)  # Increase the maximum number of iterations

# Display summary results
print(result.summary())
```

```{}
# Convert categorical variables to dummy variables, including 'FieldType', 'RosterPosition', and 'Weather'
# Be sure to handle missing data first, if any
plays.dropna(inplace=True)  # This removes any rows with missing data
plays = pd.get_dummies(plays, columns=['FieldType', 'RosterPosition', 'Weather'], drop_first=True)

# Ensure all columns are of a numeric data type
for col in plays.columns:
    if plays[col].dtype == 'object':
        plays[col] = plays[col].astype('category').cat.codes  # This converts object types to categorical

# Define the dependent and independent variables
y = plays['InjuryOccurred']
X = plays.drop('InjuryOccurred', axis=1)

# Add a constant to the independent variable set for the intercept
X = sm.add_constant(X)

# Convert X and y to numpy arrays, ensuring they are numeric
X = np.asarray(X, dtype=float)
y = np.asarray(y, dtype=float)

# Perform logistic regression
logit_model = sm.Logit(y, X)
result = logit_model.fit()

# Display summary results
print(result.summary())
```
